
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright (c) 2003-2012 by University of Queensland
% http://www.uq.edu.au
%
% Primary Business: Queensland, Australia
% Licensed under the Open Software License version 3.0
% http://www.opensource.org/licenses/osl-3.0.php
%
% Development until 2012 by Earth Systems Science Computational Center (ESSCC)
% Development since 2012 by School of Earth Sciences
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Gravity Inversion}\label{sec:forward gravity}
For the magnetic inversion we use the anomaly of the gravity acceleration~\index{gravity acceleration} of the Earth.
The controlling material parameter
is the density~\index{density} $\rho$ of the rock. 
If the density field $\rho$ is known the gravitational potential $\psi$ is given
as the solution of the PDE 
\begin{equation}\label{ref:GRAV:EQU:100}
-\psi_{,ii} = -4\pi G \cdot  \rho
\end{equation}
where $G=6.67300 \cdot 10^{-11}  \frac{m^3}{kg \cdot s^2}$ is the gravitational constant.  The gravitational potential is set to zero 
at the top of the domain $\Gamma_{top}$. On all other faces the normal component of the gravity acceleration anomaly $g_i$
is set to zero, i.e. $n_i \psi_{,i}  = 0$ with outer normal field $n_i$).
The gravity force $g_i$ is given
as the negative of the gradient of the gravity potential $\psi$:
\begin{equation}\label{ref:GRAV:EQU:101}
 g_i = - \psi_{,i} 
\end{equation} 
From the gravitational potential we can calculate the gravity acceleration anomaly via equation~(\ref{ref:GRAV:EQU:101}) to
calculate the defect to the given data. If $g^{(s)}_i$ is a measurement of the gravity acceleration anomaly for 
survey $s$ and $\omega^{(s)}_i$ is a weighting factor the data defect $J^{grav}(k)$ in the notation of Chapter~\ref{Chp:ref:introduction}
is given as
\begin{equation}\label{ref:GRAV:EQU:9}
J^{grav}(k) = \frac{\alpha}{2}\sum_{s} \int_{\Omega} ( \omega^{(s)}_i \cdot (g_{i}- g^{(s)}_i) ) ^2 dx
\end{equation} 
Summation over $i$ is performed. $\alpha$ defines a scaling factor which is typically set to one.
The cost function kernel\index{cost function!kernel} is given as
\begin{equation}\label{ref:GRAV:EQU:10}
K^{grav}(\psi_{,i},k) = \frac{\alpha}{2}\sum_{s} ( \omega^{(s)}_i \cdot (\psi_{,i}+ g^{(s)}_i) ) ^2
\end{equation} 
In practice the gravity acceleration $g^{(s)}$ is measured in vertical direction $z$ with a 
standard error deviation $\sigma^{(s)}$ at certain locations in the domain. In this case
one sets the weighting factors $\omega^{(s)}$ as
\begin{equation}\label{ref:GRAV:EQU:11}
\omega^{(s)}_i 
= \left\{
\begin{array}{lcl}
f \cdot  \frac{\delta_{iz}}{\sigma^{(s)}} & & \mbox{data are available} \\
& \mbox{ where } & \\
0 & & \mbox{ otherwise } \\
\end{array}
\right.
\end{equation} 
The scaling factor $f$ is chosen in the way that 
\begin{equation}\label{ref:GRAV:EQU:12}
\sum_{s} \int_{\Omega} ( \omega^{(s)}_i g^{(s)}_i ) ^2 dx =1 
\end{equation} 

\subsection{Usage}

\LG{Add example}

\begin{classdesc}{GravityModel}{domain, 
w, g,
\optional{, useSphericalCoordinates=False
\optional{, scale=1
\optional{, tol=1e-8}}}}
opens a gravity forward model over the \Domain \var{domain} with 
weighting factors \var{w} ($=\omega^{(s)}$) and measured gravity acceleration anomalies\var{g} ($=g^{(s)}$).
The weighting factors and the  measured gravity acceleration anomalies\var must be vectors.
If \var{useSphericalCoordinates} is \True spherical coordinates are used. 
\var{scale} sets the scaling factor $\alpha$. 
\var{tol} set the tolerance for the solution of the PDE~(\ref{ref:GRAV:EQU:100}).
The weighting factors are rescaled such that conditon~(\ref{ref:GRAV:EQU:12}) holds.
\end{classdesc}

\subsection{Gradient Calculation}
The subsection briefly explains how the gradient $\frac{\partial J^{grav}}{\partial \rho}$  of the cost function $J^{grav}$ with respect to the 
density $\rho$ is calculated. 
The magnetic potential $\psi$ from PDE~(\ref{ref:GRAV:EQU:100})  is solved in weak form:
\begin{equation}\label{ref:GRAV:EQU:201}
\int_{\Omega} q_{,i} \psi_{,i} \; dx  = - \int_{\Omega}  4\pi G \cdot q \rho\; dx 
\end{equation} 
for all $q$ with $q=0$ on $\Gamma_{top}$. In the following we set $\Psi[\cdot]=\psi$ for a given density $\cdot$ as solution
of the variational problem~(\ref{ref:GRAV:EQU:201}).If $\Gamma_{\rho}$ denotes the region of the domain where the density is known
and for a given direction $p$ with $p=0$ on $\Gamma_{\rho}$ one has
\begin{equation}\label{ref:GRAV:EQU:201}
\int_{\Omega}   \frac{\partial J^{grav}}{\partial \rho} \cdot p \; dx  = \alpha \cdot  \int_{\Omega}  
\sum_{s} (\omega^{(s)}_j \cdot 
(g^{(s)}_j-g_{j}) ) \cdot ( \omega^{(s)}_i \Psi[p]_{,i})  \; dx  
\end{equation} 
With 
\begin{equation}\label{ref:GRAV:EQU:202c}
Y_i[\psi]= \alpha \cdot \sum_{s} (\omega^{(s)}_j \cdot 
(g^{(s)}_j-g_{j}) ) \cdot  \omega^{(s)}_i
\end{equation} 
This is written as 
\begin{equation}\label{ref:GRAV:EQU:201}
\int_{\Omega}   \frac{\partial J^{grav}}{\partial \rho} \cdot p \;  dx  = \int_{\Omega}  
Y_i[\psi] \Psi[p]_{,i} \; dx  
\end{equation} 
We then set $Y^*[\psi]$ as the solution of the equation 
\begin{equation}\label{ref:GRAV:EQU:202d}
\int_{\Omega} r_{,i} Y^*[\psi]_{,i} \; dx  =  \int_{\Omega} r_{,i} ,Y_i[\psi]  \; dx  \mbox{ for all } p \mbox{ with } r=0 \mbox{ on } \Gamma_{top}
\end{equation} 
with $Y^*[\psi]=0$ on $\Gamma_{top}$. With $r=\Psi[p]$ we get
\begin{equation}\label{ref:GRAV:EQU:202d}
\int_{\Omega} \Psi[p]_{,i} Y^*[\psi]_{,i} \; dx  =  \int_{\Omega} \Psi[p]_{,i} ,Y_i[\psi]  \; dx
\end{equation} 
and from equation~(\ref{ref:GRAV:EQU:201}) with $q=Y^*[\psi]$ we get
\begin{equation}\label{ref:GRAV:EQU:20e}
\int_{\Omega} Y^*[\psi]_{,i}  \Psi[p]_{,i} \; dx  = - \int_{\Omega}  4\pi G \cdot Y^*[\psi] \cdot  p\;  dx  
\end{equation}
which leads to 
\begin{equation}\label{ref:GRAV:EQU:20ee}
\int_{\Omega} \Psi[p]_{,i} ,Y_i[\psi]  \; dx  = - \int_{\Omega}  4\pi G \cdot Y^*[\psi] \cdot  p \; dx  
\end{equation}
and finally
\begin{equation}\label{ref:GRAV:EQU:201a}
\int_{\Omega}   \frac{\partial J^{grav}}{\partial \rho} \cdot p \;  dx  = - \int_{\Omega}  
4\pi G \cdot Y^*[\psi] \cdot  p \; dx  
\end{equation} 
or 
\begin{equation}\label{ref:GRAV:EQU:201b}
\frac{\partial J^{grav}}{\partial \rho}  =- 4\pi G \cdot Y^*[\psi]
\end{equation} 

