%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright (c) 2003-2014 by University of Queensland
% http://www.uq.edu.au
%
% Primary Business: Queensland, Australia
% Licensed under the Open Software License version 3.0
% http://www.opensource.org/licenses/osl-3.0.php
%
% Development until 2012 by Earth Systems Science Computational Center (ESSCC)
% Development 2012-2013 by School of Earth Sciences
% Development from 2014 by Centre for Geoscience Computing (GeoComp)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{esysdoc}
\usepackage{makeidx} %already in the esysdoc class but putting it here prompts kile to run makeidx

% grab the handy definitions and \usepackage statements etc
\input{user_defs}
\usepackage{upquote} %used to precent LaTeX from mucking up '

% \newcommand{\hackscore}{_}

% title, author, etc stuff
\title{\emph{esys-Escript} User's Guide:\\ Solving Partial Differential Equations\\ with Escript and Finley}
\author{Lutz Gross\etal~(Editor)}
\authoraddress{
Centre for Geoscience Computing (GeoComp) \\
School of Earth Sciences \\
The University of Queensland \\
Brisbane, Australia \\
%Email: \email{esys@esscc.uq.edu.au}
}
\date{\today}
\release{3.4.2}
% \release{development}

\makeindex

% the actual start of the document
\begin{document}

\maketitle

% This must be after the make title or latex markup will end up in the title
\ifpdf
\pdfinfo{
/Author (Lutz Gross et al. (Editor))
/Title (esys-Escript User's Guide)
/Keywords (escript, PDEs)
}
\fi

\input{copyrights}


\begin{abstract}\pdfbookmark[0]{Abstract}{abstract}%
\escript is a python-based environment for implementing mathematical models, in particular those based on coupled, non-linear, time-dependent partial differential equations.



It consists of five major components
\begin{itemize}
\item \escript core library
\item finite element solver \finley (which uses fast vendor-supplied solvers or our paso linear solver library)
\item the meshing interface \pycad
\item a model library.
\item an inversion library.
\end{itemize}
The current version supports parallelization through both MPI for distributed
memory and OpenMP for shared memory.

In this release there are a number of small changes which are not backwards compatible.
Please see Appendix~\ref{app:changes} to see if your scripts will be affected.

If you use this software in your research, then we would appreciate (but do not require) a citation.
Some relevant references can be found in Appendix~\ref{app:ourrefs}.

For Python 3 support, see Appendix~\ref{app:py3}.
\end{abstract}
\input{currentdev}
\cleardoublepage\pdfbookmark[0]{Contents}{contents}%
\tableofcontents

\include{TutorialPDE}
\include{execute}
\include{escript}
\include{linearPDE}
\include{pycad}
\include{Models}
\include{finley}
\include{weipa}
\include{symbolic}

\esysappendix %So hyperef builds table of contents links properly 

%\input{modelframe}
\include{appendix}

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{\indexname}
\printindex

\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{plain}
\bibliography{esys}

\end{document}

