\chapter{The module \pyvisi}
\label{PYVISI CHAP}

\declaremodule{extension}{pyvisi}
\modulesynopsis{Python visualization interface}

\pyvisi provides an easy to use interface to the \VTK visualization 
tool. \pyvisi provides the following modules:

\begin{itemize}
\item \Scene: Displays a scene in which objects are to be rendered on.
\item \DataCollector: Deals with the source of data for visualization.
\item \Map: Displays a scalar field using a domain surface.
\item \MapOnPlaneCut: Displays a scalar field using a domain surface cut on a plane. 
\item \MapOnPlaneClip: Displays a scalar field using a domain surface clipped 
		on a plane.
\item \MapOnScalarClip: Displays a scalar field using a domain surface clipped 
		using a scalar value.
\item \Velocity: Displays a vector field using arrows.
\item \VelocityOnPlaneCut: Displays a vector field using arrows cut on a plane.
\item \VelocityOnPlaneClip: Displays a vector field using arrows clipped on a 
		plane.
\item \Ellipsoid: Displays a tensor field using spheres.
\item \EllipsoidOnPlaneCut: Displays a tensor field using spheres cut on a
        plane.
\item \EllipsoidOnPlaneClip: Displays a tensor field using spheres clipped 
        on a plane.
        
\item \Contour: Shows a scalar field by contour surfaces. 
\item \ContourOnPlane: Shows a scalar field by contour surfaces on 
a given plane.
\item \ContourOnClip: Shows a scalar field by contour surfaces on 
a given clip.

\item \Image: Displays an image.
\item \Text: Shows some 2D text.
\item \Camera: Controls the camera manipulation. 
\item \Light: Controls the light manipulation.
\item \IsoSurface: Shows a scalar field for a given value by 
an isosurface.
\item \IsoSurfaceOnPlane: Shows a scalar field for a given value by 
an isosurfaceon a given plane.
\item \IsoSurfaceOnClip: Shows a scalar field for a given vlaue by 
an isosurface on a given clip.
\item \StreamLines: Shows the path of particles in a vector field.
\item \Carpet: Shows a scalar field as plane deformated along 
the plane normal.
\item \Position: Defines the x,y and z coordinates rendered object.
\item \Transform: Defines the orientation of rendered object.
\item \Style: Defines the style of text.
\item \BlueToRed: Defines a map spectrum from blue to red.
\item \RedToBlue: Defines a map spectrum from red to blue.
\item \Plane: Defines the cutting/clipping of rendered objects.
\end{itemize}

