%!TEX root = inversion.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Copyright (c) 2003-2018 by The University of Queensland
% http://www.uq.edu.au
%
% Primary Business: Queensland, Australia
% Licensed under the Apache License, version 2.0
% http://www.apache.org/licenses/LICENSE-2.0
%
% Development until 2012 by Earth Systems Science Computational Center (ESSCC)
% Development 2012-2013 by School of Earth Sciences
% Development from 2014 by Centre for Geoscience Computing (GeoComp)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{DC resistivity inversion: 3D}\label{sec:forward DCRES}
This section will discuss DC resistivity\index{DC forward} forward modelling, as well as an \escript
class which allows for solutions of these forward problems. The DC resistivity 
forward problem is modelled via the application of Ohm's Law to the flow of current
through the ground. When sources are treated as a point sources and Ohm's Law 
is written in terms of the potential field, the equation becomes:
\begin{equation} \label{ref:dcres:eq1}
\nabla \cdot (\sigma \nabla \phi) = -I \delta(x-x_s) \delta(y-y_s) \delta(z-z_s)
\end{equation}
Where $(x,y,z)$ and $(x_s, y_s, z_s)$ are the coordinates of the observation and source
points respectively. The total potential, $\phi$, is split into primary and secondary 
potentials $\phi = \phi_p + \phi_s$, where the primary potential is analytically calculated 
as a flat half-space background model with conductivity of $\sigma_p$. 
The secondary potential is due to conductivity deviations 
from the background model and has its conductivity denoted as $\sigma_s$. 
This approach effectively removes the singularities of the Dirac delta 
source and provides more accurate results \cite{rucker2006three}.
An analytical solution is available for the primary potential of a uniform half-space due to a single pole source and is given by:
\begin{equation} \label{ref:dcres:eq2}
\phi_p = \frac{I}{2 \pi \sigma_1 R}
\end{equation}
Where $I$ is the current and $R$ is the distance from the observation points to the source.
In \escript the observation points are the nodes of the domain and $R$ is given by
\begin{equation} \label{ref:dcres:eq3}
R = \sqrt{(x-x_s)^2+(y-y_s)^2 + z^2}
\end{equation}
The secondary potential, $\phi_s$, is given by
\begin{equation}\label{ref:dcres:eq4}
-\mathbf{\nabla}\cdot\left(\sigma\,\nabla \phi_s \right)  = 
 \mathbf{\nabla}\cdot\left( \left(\sigma_p-\sigma\right)\,\nabla \phi_p  \right)
\end{equation} 
where $\sigma_p$ is the conductivity of the background half-space.
The weak form of above PDE is given by multiplication of a suitable test function, $w$, and integrating over the domain $\Omega$:
\begin{multline}\label{ref:dcres:eq5}
-\int_{\partial\Omega} \sigma\,\nabla \phi_s  \cdot \hat{n} w\,ds +
 \int_{\Omega} \sigma\,\nabla \phi_s  \cdot \nabla w\,d\Omega =\\
-\int_{\partial\Omega} \left(\sigma_p-\sigma\right)\,\nabla \phi_p  
\cdot \hat{n} w\,ds + \int_{\Omega} \left(\sigma_p-\sigma\right)\,\nabla \phi_p  \cdot \nabla w\,d\Omega 
\end{multline}
The integrals over the domain boundary provide the boundary conditions which are
implemented as Dirichlet conditions (i.e. zero potential) at all interfaces except the
top, where Neumann conditions apply (i.e. no current flux through the air-earth interface).
From the integrals over the domain, the \escript coefficients can be deduced: the 
left-hand-side conforms to \escript coefficient $A$, whereas the right-hand-side agrees
with the coefficient $X$ (see User Guide).

A number a of different configurations for electrode set-up are available \cite[pg 5]{LOKE2014}.
An \escript class is provided for each of the following survey types:
\begin{itemize}
\item Wenner alpha
\item Pole-Pole
\item Dipole-Dipole
\item Pole-Dipole
\item Schlumberger
\end{itemize}

These configurations are comprised of at least one pair of current and potential
electrodes separated by a distance $a$. In those configurations which use $n$,
electrodes in the currently active set may be separated by $na$. In the classes
that follow, the specified value of $n$ is an upper limit. That is $n$ will
start at 1 and iterate up to the value specified.

\subsection{Usage}
The DC resistivity forward modelling classes are specified as follows:

\begin{classdesc}{WennerSurvey}{self, domain, primaryConductivity, secondaryConductivity,
current, a, midPoint, directionVector, numElectrodes}
\end{classdesc}

\begin{classdesc}{polepoleSurvey}{domain, primaryConductivity, secondaryConductivity, 
current, a, midPoint, directionVector, numElectrodes}
\end{classdesc}

\begin{classdesc}{DipoleDipoleSurvey}{self, domain, primaryConductivity, secondaryConductivity,
current, a, n, midPoint, directionVector, numElectrodes}
\end{classdesc}

\begin{classdesc}{PoleDipoleSurvey}{self, domain, primaryConductivity, secondaryConductivity,
current, a, n, midPoint, directionVector, numElectrodes}
\end{classdesc}

\begin{classdesc}{SchlumbergerSurvey}{self, domain, primaryConductivity, secondaryConductivity,
current, a, n, midPoint, directionVector, numElectrodes}
\end{classdesc}

\noindent Where:
\begin{itemize}
\item \texttt{domain} is the domain which represent the half-space of interest. 
it is important that a node exists at the points where the electrodes will be placed.
\item \texttt{primaryConductivity} is a data object which defines the primary conductivity
it should be defined on the ContinuousFunction function space.
\item \texttt{secondaryConductivity} is a data object which defines the secondary conductivity
it should be defined on the ContinuousFunction function space.
\item \texttt{current} is the value of the injection current to be used in amps this is a currently a
constant.
\item \texttt{a} is the electrode separation distance.
\item \texttt{n} is the electrode separation distance multiplier.
\item \texttt{midpoint} is the centre of the survey. Electrodes will spread from this point
in the direction defined by the direction vector and in the opposite direction, placing
half of the electrodes on either side.
\item \texttt{directionVector} defines as the direction in which electrodes are spread.
\item \texttt{numElectrodes} is the number of electrodes to be used in the survey.
\end{itemize} 

When calculating the potentials the survey is moved along the set of electrodes.
The process of moving the electrodes along is repeated for each consecutive value of $n$.
As $n$ increases less potentials are calculated, this is because a greater spacing is
required and hence some electrodes are skipped. The process of building up these
pseudo-sections is covered in greater depth by Loke (2014)\cite[pg 19]{LOKE2014}.
These classes all share common member functions described below. For the surveys
where $n$ is not specified only one list will be returned. 

\begin{methoddesc}[]{getPotential}{}
Returns 3 lists, each made up of a number of lists containing primary, secondary and total
potential differences. Each of the lists contains $n$ sublists.
\end{methoddesc}

\begin{methoddesc}[]{getElectrodes}{}
Returns a list containing the positions of the electrodes
\end{methoddesc}

\begin{methoddesc}[]{getApparentResistivityPrimary}{}
Returns a series of lists containing primary apparent resistivities one for each 
value of $n$.
\end{methoddesc}

\begin{methoddesc}[]{getApparentResistivitySecondary}{}
Returns a series of lists containing secondary apparent resistivities one for each 
value of $n$.
\end{methoddesc}

\begin{methoddesc}[]{getApparentResistivityTotal}{}
Returns a series of lists containing total apparent resistivities, one for each 
value of $n$. This is generally the result of interest.
\end{methoddesc}

The apparent resistivities are calculated by applying a geometric factor to the
measured potentials.

