%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Copyright (c) 2003-2008 by University of Queensland
% Earth Systems Science Computational Center (ESSCC)
% http://www.uq.edu.au/esscc
%
% Primary Business: Queensland, Australia
% Licensed under the Open Software License version 3.0
% http://www.opensource.org/licenses/osl-3.0.php
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Linux binary installation}
\label{sec:binlinux}

\esfinley can be installed as a stand-alone bundle, containing all the required dependencies.
Alternatively, if we have a package for your distribution you can use the standard tools to install.
Please note, however, that the current binary packages do not support \openmp\footnote{This is due to a bug related to gcc 4.3.2.} or \mpi\footnote{Producing packages for \mpi requires knowing something about your computer's configuration.}.
If you need these features you may need to compile \esfinley from source (see Sections~\ref{sec:compilesrc} and \ref{sec:compileescriptlinux}.)

For more information on using the \filename{escript} command please see the User Guide.

If you are using Debian (5.0 - ``Lenny'') or Ubuntu (8.10-``Intrepid Ibex'', 9.04-``Jaunty Jackalope'') then see \Sec{sec:debian}.
For other Linux distributions refer to \Sec{sec:standalonelinux} 

\subsection{Debian 5.0(``Lenny''), Ubuntu 8.10(``Intrepid Ibex'') or 9.04(``Jaunty Jakalope'')}\label{sec:debian}

Download the \filename{escript.deb} file.
(At the time of writing we only produce deb's for the i386 architecture.)
Execute the following commands as root (you need to be in the directory containing the file).
\begin{shellCode}
 dpkg --unpack escript.deb
 aptitude install escript
\end{shellCode}

If you use sudo (for example on Ubuntu) this would be:
\begin{shellCode}
sudo dpkg --unpack escript.deb
sudo aptitude install escript
\end{shellCode}

This should install \esfinley and its dependencies on your system.
Please notify the development team if something goes wrong.


\subsection{Stand-alone bundle}\label{sec:standalonelinux}

Download the bundle and decompress it.
\begin{shellCode}
tar -xjf escript.tar.bz2 
\end{shellCode}
This will produce a directory called \filename{stand}. 
You can rename or move it as is convienient to you.
Test your installation by running:
\begin{shellCode}
 stand/escript.d/bin/escript
\end{shellCode}
You should get a normal python shell.
If you wish to save on typing you can add \filename{x/escript.d/bin} to your PATH variable (where x is the absolute path to your install).

