
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Copyright (c) 2003-2010 by University of Queensland
% Earth Systems Science Computational Center (ESSCC)
% http://www.uq.edu.au/esscc
%
% Primary Business: Queensland, Australia
% Licensed under the Open Software License version 3.0
% http://www.opensource.org/licenses/osl-3.0.php
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Darcy Flux}
\label{DARCY FLUX}
We want to calculate the velocity $u$ and pressure $p$ on a domain $\Omega$
solving the Darcy flux problem\index{Darcy flux}\index{Darcy flow}
\begin{equation}\label{DARCY PROBLEM}
\begin{array}{rcl}
u_{i} + \kappa_{ij} p_{,j} & = & g_{i} \\
u_{k,k} & = & f
\end{array}
\end{equation} 
with the boundary conditions
\begin{equation}\label{DARCY BOUNDARY}
\begin{array}{rcl}
u_{i} \; n_{i}  = u^{N}_{i}  \; n_{i} & \mbox{ on } & \Gamma_{N} \\
p = p^{D} &  \mbox{ on } & \Gamma_{D} \\ 
\end{array}
\end{equation} 
where $\Gamma_{N}$ and $\Gamma_{D}$ are a partition of the boundary of
$\Omega$ with $\Gamma_{D}$ non-empty, $n_{i}$ is the outer normal field of the
boundary of $\Omega$, $u^{N}_{i}$ and $p^{D}$ are given functions on $\Omega$,
$g_{i}$ and $f$ are given source terms and $\kappa_{ij}$ is the given
permeability.
We assume that $\kappa_{ij}$ is symmetric (which is not really required) and
positive definite, i.e. there are positive constants $\alpha_{0}$ and
$\alpha_{1}$ which are independent from the location in $\Omega$ such that
\begin{equation}
\alpha_{0} \; x_{i} x_{i} \le \kappa_{ij} x_{i} x_{j} \le \alpha_{1} \; x_{i} x_{i}
\end{equation}
for all $x_{i}$.

\subsection{Solution Method \label{DARCY SOLVE}}
It is useful to write \eqn{DARCY PROBLEM} in operator form.
For any pressure $p$ we set
\begin{equation}
(Gp)_{i} =  p_{,j}
\end{equation}
and for velocity $v$ we set
\begin{equation}
Dv = v_{k,k}
\end{equation}
With $K=(\kappa_{ij})$ we can write the Darcy problem~\ref{DARCY PROBLEM} as
\begin{equation}
\begin{array}{rcl}
u + K \, Gp & = & g \\
Du & = & f 
\end{array}
\end{equation} 
We solve this equation by minimising the functional
\begin{equation}\label{DARCY COST}
J(u,p):=\|K^{-\frac{1}{2}}(u + K \, G p - g)\|^2_{0} +  \|\lambda (Du-f) \|_{0}^2 
\end{equation} 
over all suitable $u$ and $p$.
In this equation we set $\|p\|^2_{0}=(p,p)_{0}$ with
\begin{equation}
(p,q)_{0} = \int_{\Omega } p\cdot q \, dx
\end{equation} 
The factor $\lambda>0$ is a weighting factor.
A simple calculation shows that one has to solve
\begin{equation}
( K^{-1} (v + K \, Gq) , u +K \, G p - g)_{0} +  (\lambda Dv,\lambda (Du-f) )_{0} =0 
\end{equation} 
for all velocities $v$ and pressure $q$ which fulfill the homogeneous boundary
conditions~\ref{DARCY BOUNDARY}.
This so-called variational equation can be translated back into operator notation
\begin{equation}
\begin{array}{rcl}
(K^{-1}+ D^*\lambda^2 D)u + Gp & = &  D^*\lambda f + K^{-1} g \\
G^*u  + G^*K \, G p & = & G^*g \\ 
\end{array}
\end{equation} 
where $D^*$ and $Q^*$ denote the adjoint operators with respect to $(.,.)_{0}$.
In~\cite{LEASTSQUARESFEM1994} it has been shown that this problem is
continuous and coercive and therefore has a unique solution.
Standard FEM methods can be used for discretization. It is also possible to
solve the problem in coupled form, however this approach leads in some cases
to a very ill-conditioned stiffness matrix, in particular in the case of a
very small or large permeability ($\alpha_{1} \ll 1$ or $\alpha_{0} \gg 1$).

The approach we are taking is to eliminate the velocity $v$ from the problem.
Assuming that $p$ is known we have\footnote{notice
that $K^{-1}+\lambda D^*D$ is coercive}
\begin{equation}\label{DARCY V FORM}
v= (K^{-1}+ D^*\lambda^2 D)^{-1} ( D^*\lambda f + K^{-1} g - Gp)
\end{equation} 
which is inserted into the second equation
\begin{equation}
G^* (K^{-1}+\lambda D^*D)^{-1} (\lambda D^*f + K^{-1} g - Gp) + G^* KG p = G^*g 
\end{equation} 
which is
\begin{equation}
G^* ( K - (K^{-1}+ D^*\lambda^2 D)^{-1} ) G p = G^* (g-(K^{-1}+D^*\lambda^2 D)^{-1} ( D^*\lambda f + K^{-1} g) ) 
\end{equation} 
We use the PCG\index{linear solver!PCG}\index{PCG} method to solve this.
The residual $r$ is given as
\begin{equation}
\begin{array}{rcl}
r & =&  G^* \left( g - K\, G p - v \right)
\end{array}
\end{equation} 
for the current pressure approximation $p$ and current velocity $v$ defined by
\eqn{DARCY V FORM}.
So in a particular implementation we use $\hat{r}=g-K\, Gp-v$ to represent the residual.
The evaluation of the iteration operator for a given $p$ is then returning
$Qp+v$ where $v$ is the solution of
\begin{equation}\label{UPDATE W}
(K^{-1}+ D^*\lambda^2 D)v = Gp
\end{equation}
To derive a preconditioner we use the identity
\begin{equation}
 \begin{array}{rcl}

G^* ( K - (K^{-1}+ D^*\lambda^2 D)^{-1} ) G  & = & G^* (I - (I + K D^*\lambda^2 D)^{-1}) K G \\
                               & \approx &  G^* (K D^*\lambda^2 D) K G  \\
			       & =  & G^*  K ( D^*  \lambda^2 D) K G \\
			       & \approx  & G^* \frac{\lambda^2}{dx^2} K^2 G
\end{array}
\end{equation} 
where $dx$ is the local mesh size and we use the approximation
\begin{equation}
D^*  \lambda^2 D \approx \frac{\lambda^2}{dx^2}
\end{equation} 
Therefore we use $G^* \frac{\lambda^2}{dx^2} K^2 G$ as a preconditioner.
To evaluate the preconditioner we need to solve the equation
\begin{equation}\label{UPDATE P}
 G^* \frac{\lambda^2}{dx^2} K^2 G p =  G^* \hat{r}
\end{equation}
It remains to answer the question how to choose $\lambda$.
We need to balance the first and second term in $J(u,p)$ in \eqn{DARCY COST}.
We inspect $J$ for $(\hat{u}, \hat{p})$ which is a perturbed exact solution $(u,p)$.
Assuming $\hat{u}=u+u_{0}e^{ik^tx}$ and $\hat{p}=p+p_{0}e^{ik^tx}$ and
constant $K$ we get
\begin{equation}
J(\hat{u},\hat{p}) = C \left[ ( \|K^{-1}\|_{2} |u_{0}|^2 + \|K\|_{2} \|k\|_{2}^2 |p_{0}|^2| )
+ \lambda^2 \|k\|_{2}^2 |u_{0}|^2  \right]
\end{equation} 
with some constant $C>0$.
The first two terms and the third term correspond to the first term and
second term in the definition of $J(u,p)$ in \eqn{DARCY COST}.
For small $\|k\|_{2}$ (i.e. for a smooth perturbation) $J(\hat{u},\hat{p})$ is
dominated by $\|K^{-1}\|_{2} |u_{0}|^2$.
To scale the second term which corresponds to the incompressibility condition
for the velocity we need to meet the condition $\|K^{-1}\|_{2} = \lambda^2 \|k\|_{2}^2$.
Taking the boundary conditions into consideration the smallest possible value
for $\|k\|_{2} = \frac{\pi}{l}$ where $l$ is the longest edge of the domain.
This leads to
\begin{equation}\label{DARCY LAMBDA}
\lambda = \|K^{-1}\|_{2}^{\frac{1}{2}} \frac{l}{\pi}
\end{equation} 
Notice that with this setting the preconditioner
$G^* \frac{\lambda^2}{dx^2} K^2 G$ becomes equivalent to $G^* K G$ if $K$ is a
diagonal matrix and the mesh has a constant mesh size.

The residual norm used in PCG is given as
\begin{equation}\label{DARCY R NORM}
\|r\|_{PCG}^2 = \int r (G^* \frac{\lambda^2}{dx^2} K^2 G)^{-1} r \; dx =\int \hat{r}  G ( G^* \frac{\lambda^2}{dx^2} K^2 G)^{-1}  G^* \hat{r} \; dx \approx
\int \hat{r} K^{-1}  \hat{r} \; dx 
\end{equation}
The iteration is terminated if
\begin{equation}\label{DARCY STOP}
\|r\|_{PCG} \le \mbox{ATOL}
\end{equation}
where we set
\begin{equation}\label{DARCY ATOL DEF}
\mbox{ATOL} = \mbox{atol} + \mbox{rtol} \cdot \left(\frac{1}{\|K^{-\frac{1}{2}}v\|_{0}} + \frac{1}{\|K^{\frac{1}{2}} G p\|_{0}} \right)^{-1} 
\end{equation}
where rtol is a given relative tolerance and atol is a given absolute
tolerance (typically $=0$).
Notice that if $Gp$ and $v$ both are zero, the pair $(0,p)$ is a solution.
The problem is that ATOL is depending on the solution $p$ and $v$ calculated
form~\ref{DARCY V FORM}.
In practice one can use the initial guess for $p$ to get a first value for ATOL.
If the stopping criterion is met in the PCG iteration, a new $v$ is calculated
from the current pressure approximation and ATOL is recalculated.
If \ref{DARCY STOP} is still fulfilled the calculation is terminated and
$(v,p)$ is returned. Otherwise PCG is restarted with a new ATOL.

\subsection{Functions}
\begin{classdesc}{DarcyFlow}{domain}
opens the Darcy flux problem\index{Darcy flux} on the \Domain domain.
\end{classdesc}

\begin{methoddesc}[DarcyFlow]{setValue}{\optional{f=None, \optional{g=None, \optional{location_of_fixed_pressure=None, \optional{location_of_fixed_flux=None, 
\\\optional{permeability=None}}}}}}
assigns values to the model parameters. Values can be assigned using various
calls -- in particular in a time dependent problem only values that change
over time need to be reset. The permeability can be defined as a scalar
(isotropic), a vector (orthotropic) or a matrix (anisotropic).
\var{f} and \var{g} are the corresponding parameters in~\ref{DARCY PROBLEM}.
The locations and components where the flux is prescribed are set by positive
values in \var{location_of_fixed_flux}.
The locations where the pressure is prescribed are set by by positive values
of \var{location_of_fixed_pressure}.
The values of the pressure and flux are defined by the initial guess.
Notice that at any point on the boundary of the domain the pressure or the
normal component of the flux must be defined. There must be at least one point
where the pressure is prescribed.
The method will try to cast the given values to appropriate \Data class objects.
\end{methoddesc}

\begin{methoddesc}[DarcyFlow]{setTolerance}{\optional{rtol=1e-4}}
sets the relative tolerance \var{rtol} in \ref{DARCY ATOL DEF}.
\end{methoddesc}

\begin{methoddesc}[DarcyFlow]{setAbsoluteTolerance}{\optional{atol=0.}}
sets the absolute tolerance \var{atol} in \ref{DARCY ATOL DEF}.
\end{methoddesc}

\begin{methoddesc}[DarcyFlow]{getSolverOptionsFlux}{}
returns the solver options used to solve the flux problems~(\ref{DARCY V FORM}) and~(\ref{UPDATE W}).
Use this \SolverOptions object to control the solution algorithms.
\end{methoddesc}

\begin{methoddesc}[DarcyFlow]{getSolverOptionsPressure}{}
returns a \SolverOptions object with the options used to solve the pressure
problem~(\ref{UPDATE P}) as a preconditioner.
Use this object to control the solution algorithms.
\end{methoddesc}

\begin{methoddesc}[DarcyFlow]{solve}{u0,p0, \optional{max_iter=100, \optional{verbose=False}}}
solves the problem and returns approximations for the flux $v$ and the pressure $p$.
\var{u0} and \var{p0} define initial guesses for flux and pressure.
Values marked by positive values \var{location_of_fixed_flux} and
\var{location_of_fixed_pressure}, respectively, are kept unchanged.
\var{max_iter} sets the maximum number of iteration steps allowed for solving
the coupled problem.
\end{methoddesc}

%\subsection{Example: Gravity Flow}
%later

